Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-28 paper/ijcai16.tex
descrizione dei metodi bayesiani

Preferences \cite{Peintner2008} play an important role in artificial intelligence,
including fields spanning from recommender systems to automatic planning, 
from non monotonic reasoning to computational social choice and algorithmic decision theory.
The task of eliciting or learning preferences is therefore a crucial; typically only limited information about the user's preferences will be available and the cost of obtaining additional preference information will be usually high (from a cognitive and/or computational point of view).
The goal in preference elicitation is that of gaining enough information about the user's preferences (typically represented by an utility function $u$, unknown to the system) in order to recommend a decision or a course of action.

The assessment of preferences have received considerable attention in the literature, with early works confined to the operation research and management literature mostly \cite{}.
In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective; within the machine learning field the term preference learning has been catching up.
%While preferences are often assessed using some ad-hoc rules \cite{},
Recently, in the last decade or so, a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas2007,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.

@@

Preference elicitation in configuration problems has been previously tackled with regret-based elicitation \cite{boutilier2006,braziunas2007}, where minimax regret is used both
as a robust recommendation criterion and as a technique to drive elicitation in configuration problems.
The main limitation of the minimax regret approach is the lack of tolerance with respect to user inconsistency (the user might state that item x is preferred to y, even though the latter has higher utility).

@@

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=16em]{figures/setmargin}
    \end{center}
    \caption{\label{fig:setmargin} The intuition behind setwise max-margin; the red points are utility vectors $\vw^{i}$ $\vw^{j}$, the red line corresponds to the constraint $\langle \vw, \vx^{1} - \vx^{2} \rangle \ge 0$.}
\end{figure}

@@

\paragraph{Set-wise max-margin.} The full {\sc SetMargin} algorithm
follows the usual preference elicitation loop. Starting from an
initially empty set of user responses $\calD$, it repeatedly solves
the MILP problem above using the set $\calD$ to enforce ranking
constraints on the weight vectors $\{\vw^i\}$. The generated
configurations $\{\vx^i\}$, which are chosen to be as good as possible
with respect to the estimated user preferences, and as diverse as
possible, are then employed to formulate a set of user queries. The
new replies are added to $\calD$ and the whole procedure is
repeated. The algorithm terminates after a fixed number of iterations,
or once another suitable condition is met. 
%\andrea{@Paolo: what termination condition for pref elicitation?}  
Termination might be left to the user to decide (see for example \ref{} for discussion on user-involved recommender systems); the system can however propose
to terminate when the difference between the utility weights is small enough.

@@

We compare {\sc SetMargin} against two strong Bayesian baselines:
\cite{guo2010real} and \cite{viappiani2010optimal}.
\stefano{Paolo, could you describe the baselines?}

We adopt the {\em indifference-augmented} Bradley-Terry user response
model introduced in~\cite{guo2010real}. According to the classical
(without indifference) Bradley-Terry model~\cite{BraTer52}, when queried about a pair
of configurations, the probability that a user will prefer
configuration $\vx^i$ over the alternative $\vx^j$ amounts to:
%
$$ (1 + \exp(-\lambda_1 \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
%
where $\vw$ is the weight vector of the true underlying user utility.
In other words, the probability grows with the difference between the
utilities of the two items. Support for indifference has been
introduced in~\cite{guo2010real} by postulating that the probability
of the user being indifferent about the two configurations depends on
how close their utilities are, as in:
%
$$ \exp(\lambda_2 |\langle\vw,\vx^i - \vx^j\rangle|) $$
%
The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
simulations, as in~\cite{guo2010real}.

@@

Users were simulated by drawing $20$ random utility vectors from each of four
different distributions. The first two mimic those used
in~\cite{guo2010real}\footnote{The normal distribution reported
in~\cite{guo2010real} differs slightly from the one implemented in the code of
their experimental setting. 
In our experiments we use the latter.}:
(1) a uniform distribution over $[1, 100]$ for each individual weight, and (2)
a normal distribution with mean $25$ and %covariance
standard deviation $\frac{25}{3}$ (each attribute is sample i.i.d).  
We further produced two novel
{\em sparse} sets of weight vectors by setting to zero $80\%$ of the entries of the
vectors sampled in the uniform and normal settings uniformly at random. 
All methods were evaluated on the very same weight vectors. All users had
a budget of 100 iterations; for {\sc SetMargin} we used a $0.01$
threshold on the utility loss for early stopping (that does not affect the
cross-validation procedure). \stefano{I have no idea if the other methods
have any kind of early stopping logic.}

@@

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions. Solution quality is measured in terms of utility loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that Guo and Viappiani (both QI and
EUS) are single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a 2.8 GHz Intel
Xeon CPU with 8 cores and 32 GiB of RAM.
In all figures, thick lines indicate median values over the different
users, shaded areas indicate standard deviations. Our {\sc SetMargin}
algorithm for $k=2$ is reported in orange, Guo in blue, Viappiani QI
in green and Viappiani EUS in gray. For all algorithms, one iteration
corresponds to a single pairwise query to the user.  Results for $r=3$
and $r=4$ are reported.  For dense weight vector distributions (first
two rows), our approach achieves results which are indistinguishable
from the competitors in a fraction of their time.  Indeed, all
Bayesian approaches become quickly intractable for growing values of $r$,
%cannot scale over $r=4$ (for $r=5$, Guo did not
%finish a single user after 24 hours for, Viappiani QI takes 20 minutes
%per user), \andrea{20 minuti mi paiono pochi, sei sicuro?}  
while our algorithm can easily scale to much larger datasets, as will
be shown later on. For sparse weight vector distributions (last two
rows) our approach, in addition to being substantially faster on each
iteration, requires less queries in order to reach optimal
solutions. This is an expected result as the sparsification norm in
our formulation ($\| \vw \|_1$) is enforcing sparsity in the weight
vectors, while none of the other approaches are explicitly designed to
favour sparse functions.

@@

\paolo{Maybe a plot of an experiment with increased noise in user response? (for example, $\beta=0.1$ instead of $1$.}

>>>>

Preferences \cite{Peintner2008} play an important role in artificial intelligence,
including fields spanning from recommender systems to automatic planning, 
from non monotonic reasoning to computational social choice and algorithmic decision theory.
The task of eliciting or learning preferences is therefore a crucial; typically only limited information about the user's preferences will be available and the cost of obtaining additional preference information will be high (from a cognitive and/or computational point of view).
The goal in preference elicitation is that of gaining enough information about the user's preferences %(typically represented by an utility function $u$, unknown to the system) 
in order to recommend a decision or a course of action.

The assessment of preferences have received considerable attention in the literature, with early works confined to the operation research and management literature mostly \cite{White1984,JacquetLagreze95}.
%In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective%; within the machine learning field the term preference learning has been catching up.
%While preferences are often assessed using some ad-hoc rules \cite{},
Recently, in the last decade or so, a number of artificial intelligence researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas2007,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.

@@

Preference elicitation in configuration problems has been previously tackled with regret-based elicitation \cite{boutilier2006,braziunas2007}, where minimax regret is used both
as a robust recommendation criterion and as a technique to drive elicitation in configuration problems.
The main limitation of the minimax regret approach is the lack of tolerance with respect to user inconsistency (the user might state that an item is preferred to another one, even though the latter has higher utility).

@@

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=10em]{figures/setmargin}
    \end{center}
    \caption{\label{fig:setmargin} The intuition behind setwise max-margin; the red points are utility vectors $\vw^{i}$ $\vw^{j}$, the red line corresponds to the constraint $\langle \vw, \vx^{1} - \vx^{2} \rangle \ge 0$.}
\end{figure}

@@

\paragraph{Set-wise max-margin.} The full {\sc SetMargin} algorithm
follows the usual preference elicitation loop. Starting from an
initially empty set of user responses $\calD$, it repeatedly solves
the MILP problem above using the set $\calD$ to enforce ranking
constraints on the weight vectors $\{\vw^i\}$. The generated
configurations $\{\vx^i\}$, which are chosen to be as good as possible
with respect to the estimated user preferences, and as diverse as
possible, are then employed to formulate a set of user queries. The
new replies are added to $\calD$ and the whole procedure is
repeated. The algorithm terminates after a fixed number of iterations,
or once another suitable condition is met. 
%\andrea{@Paolo: what termination condition for pref elicitation?}  
Termination might be left to the user to decide (see for example \cite{Reilly2007}); 
alternatively we can terminate when the difference between the utility weights is small enough.

@@

We compare {\sc SetMargin} against the following competitive Bayesian baselines:
% and \cite{viappiani2010optimal}.

\paolo{Se bisogno di spazio, posso accorciare la descrizione.}
\begin{itemize}
 \item The Bayesian approach from \cite{guo2010real}, selecting queries according to 
 {\em restricted uninformed VOI}, an heuristic approximation of value-of-information a query, and performing inference using  TrueSkill$^{TM}$ \cite{HerbrichMG06} (an inference scheme based on expectation propagation \cite{Minka01}).
 \item The Bayesian framework of \cite{viappiani2010optimal}, using Monte Carlo methods for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called EUS), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
 \item {\em Query Iteration} (referred as QI below), also from \cite{viappiani2010optimal}, an even faster query selection method based on sampling a set of utility vectors, selecting the items that maximize utility for each of the vector, forming a first proposal for a query set. Then, a new set is computed by considering the optimal items in the associated  beliefs, and the process is repeated until convergence.
\end{itemize}

We adopt the {\em indifference-augmented} Bradley-Terry user response
model introduced in~\cite{guo2010real}. According to the classical
(without indifference) Bradley-Terry model~\cite{BraTer52}, when queried about a pair
of configurations, the probability that a user will prefer
configuration $\vx^i$ over the alternative $\vx^j$ amounts to:
%
$$ (1 + \exp(-\lambda_1 \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
%
where $\vw$ is the weight vector of the true underlying user utility.
In other words, the probability grows with the difference between the
utilities of the two items. Support for indifference has been
introduced in~\cite{guo2010real} by postulating that the probability
of the user being indifferent about the two configurations depends on
how close their utilities are, as in:
%
$$ \exp(-\lambda_2 |\langle\vw,\vx^i - \vx^j\rangle|) $$
%
The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
simulations, as in~\cite{guo2010real}.

@@

Users were simulated by drawing $20$ random utility vectors from each of four
different distributions. The first two mimic those used
in~\cite{guo2010real}\footnote{The normal distribution reported
in~\cite{guo2010real} differs slightly from the one implemented in the code of
their experimental setting. 
In our experiments we use the latter.}:
(1) a uniform distribution over $[1, 100]$ for each individual weight, and (2)
a normal distribution with mean $25$ and %covariance
standard deviation $\frac{25}{3}$ (each attribute is sample i.i.d).  
We further produced two novel
{\em sparse} sets of weight vectors by setting to zero $80\%$ of the entries of the
vectors sampled in the uniform and normal settings uniformly at random. 
All methods were evaluated on the very same weight vectors. All users had
a budget of 100 iterations; for {\sc SetMargin} we used a $0.01$
threshold on the utility loss for early stopping (that does not affect the
cross-validation procedure).
The Bayesian methods can use the estimated loss to decide when to stop (the expectation
of the utility loss with respect to the current probabilistic belief)
\footnote{In our experiments, we left the Bayesian methods to run the full amount of iterations,
even when the estimated loss was very low.}.

@@

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions. Solution quality is measured in terms of utility loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that Guo and Viappiani (both QI and
EUS) are single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a 2.8 GHz Intel
Xeon CPU with 8 cores and 32 GiB of RAM.
In all figures, thick lines indicate median values over the different
users, shaded areas indicate standard deviations. Our {\sc SetMargin}
algorithm for $k=2$ is reported in orange, Guo in blue, Viappiani QI
in green and Viappiani EUS in gray. For all algorithms, one iteration
corresponds to a single pairwise query to the user.  Results for $r=3$
and $r=4$ are reported.  For dense weight vector distributions (first
two rows), our approach achieves results which are indistinguishable
from the competitors in a fraction of their time.  Indeed, all
Bayesian approaches become quickly intractable for growing values of $r$,
%cannot scale over $r=4$ (for $r=5$, Guo did not
%finish a single user after 24 hours for, Viappiani QI takes 20 minutes
%per user), \andrea{20 minuti mi paiono pochi, sei sicuro?}  
while our algorithm can easily scale to much larger datasets, as will
be shown later on. For sparse weight vector distributions (last two
rows) our approach, in addition to being substantially faster on each
iteration, requires less queries in order to reach optimal
solutions. This is an expected result as the sparsification norm in
our formulation ($\| \vw \|_1$) is enforcing sparsity in the weights, 
while none of the other approaches is %explicitly 
designed to favour sparse functions.

@@

%\paolo{Maybe a plot of an experiment with increased noise in user response? (for example, $\beta=0.1$ instead of $1$.}
